% !TEX root = Documentation.tex

\subsection{home.php}

The home.php page serves as the central page of the whole project, providing basic information on the use of the frontend and showing all navigation entries to refer to other pages 
within the frontend. To ensure that non permissioned users can not access pages such as \hyperref[create-user.php]{create-user.php}, these entries are just shown to users with roles 
higher than 1 or 2. The same happens for the dropdown menu, which is included in all pages, to allow easy navigation. 

\subsection{traffic.php}
The traffic page is the evaluation page to the logdb database, which keeps track of users accessing the login.php page and successful logins. As these data are highly sensitive and 
not important for the daily use of the database, the access to this page is limited to users with the role admin (role code 4). The presented data are split into different tables, 
one for the login.php page, displaying amount of accesses, time and IP address. Within that table, there is one line for all accesses, ordered by IP address corresponding to that access. Another table displays the data for the logged in users, showing IP address, time of login and user invoked. And finally there is a table displaying all login attempts, which failed due to incorrect credentials. This table should enable the recognition of brute force attacks on given accounts.\\
The different table lines rely on different SQL queries, to gain there results, even if they request the same table.

\subsection{database-update.php}
The database-update.php provides the graphical interface to import external data into the database. Therefore, it is the only way, the users of the interface can modify the database 
on a larger scale, than in the manual inputs, which are provided in the other pages. The processing of the files takes place in the following upload.php file. Similar to the 
create-user.php page, the database-update.php page can just be accessed by users with higher roles. 

\subsection{upload.php}
The upload.php script processes the files selected in the database-update.php file and pastes them into the files database table in the agri\_star\_001 database.
