% !TEX root = Documentation.tex

\section{Introduction}

The 2Org-Cows project is multinational and multi-partner project of research institutions in organic agriculture in Europe. It aims to improve the breeding of dairy cattle for 
organic dairy cow keeping and therefore aims to collect data to measure different parameters of animal welfare, health and keeping.\\
In order to analyze all these data, a huge database has been created, which consists of the following components:
\begin{itemize}
 \item a webbased, graphical user interface for the final user to access and query the database using a standard webbrowser
 \item a variety of interfaces to other programs
 \item software to analyze database stored datasets using big data methods, mainly based on R and on Python3
 \item MySQL databases as the storage backbone of the whole software and project
 \item Ubuntu 16.04 operating system, PHP7 scripting language and Apache2 webserver to provide the user with the data required
\end{itemize}
In the following all parts of the software for the 2Org-Cows project should be documented. Please note that this file is work in progress and will therefore change steadily.

\section{Common information regarding the coding}
The project is generally based, in its programming, on several concepts, which should be briefly displayed in the following. The programming is therefore differenciated in the 
parts regarded. 
\paragraph{Web Frontend}
The web frontend is largely written in PHP7 and HTML5, following the following paradigms: 
\begin{itemize}
 \item clear structure of code
 \item clear separation of display code and code for the processing of input
 \item easy to write, understand and maintain code
 \item Mostly procedural code style, only exception is the database system, which is based on object oriented code style
 \item All available standard security solutions for PHP and HTML code should be applied, to ensure secure operation of the database. This includes escaping of all strings, using 
 mysqli::real\_escape\_string() for all values, which is executed for all strings being in some kind of contact to the database, and htmlspecialchars() for all other strings. 
 The insert scripts perform type checking operations on the content to insert, to limit the risk of inconsistent data.
 \item SQL queries, including user generated content, are usually proceeded as prepared statements, to limit the risk of sql injections.
\end{itemize}
The appearance of all sites is specified in the 2Org-Cows.css, which acts as the only stylesheet in the whole user interface. Only in some limited situations, style information is 
provided in HTML5 inline commands. These cases mainly occur in table-styled content, where the font definition is separetely on each sheet.  
In the first development stage only limited responsiveness of the design is achieved, further development will focus on the responsiveness of the design
and the right alignment of user interface elements in HTML5 on small screen devices, such as smartphones. Due to the extensive size of all pages, this might lead to further 
reconstruction of the page and therefore to a special mobile version of the whole frontend.
\paragraph{Database}
The database design aims to provide a star schemed database, with as little programming and connection efforts as possible. Therefore, the overall database largely consists of 
independent tables, which are linked together by the code of the frontend and the database interfaces.